---
title: "Assignment 2: Coding Basics"
author: "Lehe, Xu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your first and last name into the file name (e.g., "FirstLast_A02_CodingBasics.Rmd") prior to submission.



## Basics Day 1

1. Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name. 

2. Compute the mean and median of this sequence. 

3. Ask R to determine whether the mean is greater than the median. 

4. Insert comments in your code to describe what you are doing.

```{r}
#1 Generate a sequence named hun_sequence, contaning numbers from one to 100, increasing by fours. 
hun_sequence <- seq(1, 100, 4)

hun_sequence

#2 Find the mean and median of hun_sequence
mean(hun_sequence)
median(hun_sequence)

#3 Determine whether the mean of hun_sequence is greater than the median of hun_sequence
mean(hun_sequence)>median(hun_sequence)

```

## Basics Day 2

5. Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6. Label each vector with a comment on what type of vector it is.

7. Combine each of the vectors into a data frame. Assign the data frame an informative name.

8. Label the columns of your data frame with informative titles.

```{r}
#5 
student_name <- c("Lily","Bily","Cici","Andy") #names of students (Nominal data)
test_score<-c(51, 55, 52, 50) # test scores out of a total 100 points (discrete data)
test_score>=50
score_df <- data.frame("Name"=student_name,"Score"=test_score)
score_df
```

9. QUESTION: How is this data frame different from a matrix?

> Answer:  In a data frame the columns contain different types of data, but in a matrix all the elements are the same type of data.Also, each row in data frame is parallel to each other; whereas in matrix, data are all independent.

10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.
```{r}
Scorejudge <- function(x){
  ifelse(x>=50, "TRUE", "FALSE") 
  
}

```

11. Apply your function to the vector with test scores that you created in number 5. 

```{r}
Score <- Scorejudge(test_score); Score
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: ifelse worked, but 'if' and 'else' not. Becuasue ifelse can be used to treat vector, but 'if' and 'else' cannot. 
